---
title: "REL NW Washington"
output: 
    html_document:
      theme: spacelab
      highlight: tango
      toc: TRUE
      toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r setup 1, include=FALSE}
library(tidyverse)
library(rio)
library(here)
library(colorblindr)
library(gghighlight)
library(forcats)
library(ggrepel)
library(gt)
library(knitr)
library(kableExtra)
library(reactable)
library(plotly)
library(SCVA)
library(patchwork)

theme_set(theme_minimal(15) +
            theme(legend.position = "bottom",
                  panel.grid.major.x = element_line(color = "gray60"),
                  panel.grid.minor.x = element_blank(),
                  panel.grid.major.y = element_blank())
          )

```

```{r, include=FALSE}
courses <- import(here("data", "courses.sav"),
               setclass = "tbl_df") %>% 
  characterize() %>% 
  janitor::clean_names() %>% 
  filter(state == "WA")

wa_courses <- import(here("data", "wa_courses.xlsx"),
               setclass = "tbl_df") %>% 
  janitor::clean_names() 

str(wa_courses)

quick_data <- wa_courses %>% 
  select(1, 2, 3, 5, 15, 46, 52) %>% 
  na.omit()

quick_data[sapply(quick_data, is.character)] <- lapply(quick_data[sapply(quick_data, is.character)], as.factor)

str(quick_data)
```

```{r, include=FALSE}
quick_data2 <- quick_data %>% 
  mutate(locale_code_4 = recode(locale_code_4,
                                "1" = "City",
                                "2" = "Suburb",
                                "3" = "Town",
                                "4" = "Rural"))
```

## Summary 

### School Type 

```{r, include=FALSE}

sch_type <- quick_data2 %>% 
  group_by(locale_code_4) %>% 
  count()

sch_type_plot <- sch_type %>% 
  ggplot(aes(fct_reorder(locale_code_4, n), n)) +
  geom_col(fill = "#56B4E9",
             alpha = 0.7) +
  geom_text(aes(locale_code_4, n, label = n),
            size = 4,
            nudge_y = -2.0,
            nudge_x = 0.3) +
  labs(x = "School Type",
       y = "Total")
```

### Advanced Course Offerings  

```{r, include=TRUE}
ggplotly(sch_type_plot)
```

```{r, include=FALSE}
str(quick_data2)

quick_data2[sapply(quick_data2, is.character)] <- lapply(quick_data2[sapply(quick_data2, is.character)],as.factor)

str(quick_data2)
```

```{r, include=FALSE}
course_offer <- quick_data2 %>% 
  group_by(locale_code_4, advm_ind) %>% 
  count()

course_offer_table <- course_offer %>% 
    reactable(
    columns = list(
      locale_code_4 = colDef(name = "School Type",
                   align = "center"),
      advm_ind = colDef(name = "Advanced Courses",
                        align = "center"),
      n = colDef(name = "Total",
                 align = "center")),
    striped = TRUE,
    outlined = TRUE,
    compact = TRUE,
    highlight = TRUE,
    bordered = TRUE
  )
```

```{r, include=TRUE}
course_offer_table
```

### *Chi-Square* Analysis

```{r, include=FALSE}

47 + 33 
# 80 rural high schools 

47 / 80

22 + 76

76 / 98

25 + 80

80 / 105

28 + 36

36/64
```

58.75% of `Rural` high schools and 56.25% of high schools classified as `Town` offer advanced courses, while 77.55% of `City` high schools and 76.19% of `Suburban` high schools offer advanced courses. 

```{r, include=FALSE}
head(quick_data2)

chi_data <- quick_data2 %>% 
  select(locale_code_4, advm_ind)

str(chi_data)

table(chi_data$locale_code_4, chi_data$advm_ind)

```

```{r, include=TRUE}
test_chi <- chisq.test(table(chi_data$locale_code_4, chi_data$advm_ind))
test_chi
```

```{r, include=TRUE}
test_chi$statistic
```

```{r, include=TRUE}
test_chi$p.value
```

```{r, include=TRUE}
summary(table(chi_data$locale_code_4, chi_data$advm_ind))
```

The results of the chi-square test are significant with a *p*-value of .002, suggesting there is a significant relationship between school type and the status of advanced course offerings. 

$$X^2 = 14.647, df = 3, p = .002$$


### *Chi-Square* Visualizations 

```{r, include=FALSE}
# install.packages("gplots")
library(gplots)
library(graphics)
# install.packages("vcd")
library(vcd)

chi_data2 <- chi_data %>% 
  rename("SchoolType" = locale_code_4,
         "AdvancedCourses" = advm_ind)

```

```{r, include=TRUE}
mosaic(~ SchoolType + AdvancedCourses,
  direction = c("v", "h"),
  data = chi_data2,
  shade = TRUE
)
```


```{r, include=FALSE}
# install.packages("ggstatsplot")
library(ggstatsplot)
```

```{r, include=TRUE}
ggbarstats(
  data = chi_data2,
  x = SchoolType,
  y = AdvancedCourses) 

```